---
title: "R Notebook"
output: html_notebook
---

### Libraries

```{r}
library(chromasc)
library(ggplot2)
library(SCORPIUS)
library(Seurat)
library(pheatmap)
```

### Load R14 cells and preprocess

```{r}
r14 <- Read10X(data.dir="../2019-03-10_c14_aggr-results/")
r14 <- CreateSeuratObject(raw.data=r14, min.cells=3, min.genes=3, project="r14")
mito.genes <- grep(pattern = "^MT-", x = rownames(x = r14@data), value = TRUE)
percent.mito <- Matrix::colSums(r14@raw.data[mito.genes, ])/Matrix::colSums(r14@raw.data)
r14 <- AddMetaData(object = r14, metadata = percent.mito, col.name = "percent.mito")
```

```{r}
r14 <- FilterCells(r14, subset.names=c("nGene", "percent.mito"), low.thresholds=c(200,-Inf),
                   high.thresholds=c(3700, 0.1))
r14 <- NormalizeData(object=r14, normalization.method="LogNormalize", scale.factor=10000)
r14 <- ScaleData(object=r14)
```

### Load gene sets

```{r}
hallmarks <- chromasc::load_gene_sets("../2019-03_cc-time/genesets/h.all.v6.2.symbols.gmt")
```

### Compute null parameters

```{r}
null_params <- chromasc::compute_null_params(r14, hallmarks, n_random_cells=50)
```

### Build pseudotime trajectory with SCORPIUS

```{r}
r14@data <- as.matrix(r14@data)
```

### Load Seurat cell cycle genes

```{r}
cc.genes <- readLines(con="../regev_lab_cell_cycle_genes.txt")
s.genes <- cc.genes[1:43]
g2m.genes <- cc.genes[44:97]
r14 <- CellCycleScoring(object=r14, s.genes=s.genes, g2m.genes=g2m.genes, set.ident=T)
```

### Convert sc data to matrix and keep only cell cycle genes

```{r}
common.genes <- intersect(rownames(r14@scale.data), cc.genes)
r14.mat <- t(r14@scale.data[common.genes,])
```

### Reduce dimensionality with SCORPIUS

```{r}
space <- reduce_dimensionality(r14.mat, correlation_distance, ndim=3)
```

### Infer trajectory using SCORPIUS

```{r}
traj <- infer_trajectory(space)
```

### Draw trajectory plot

```{r}
draw_trajectory_plot(
  space, 
  progression_group = r14@ident,
  path = traj$path,
  contour = TRUE
)
```

```{r}
gimp <- gene_importances(r14.mat, traj$time, num_permutations = 0, num_threads = 16)
gene_sel <- gimp[1:50,]
expr_sel <- r14.mat[,gene_sel$gene]
```

```{r}
draw_trajectory_heatmap(expr_sel, traj$time, r14@ident)
```

### Parse metadata for line plotting

```{r}
exp.names <- c("r14_10u", "s02", "r14_15u_E", "s03", "r14_5u",
               "r14_15u_L", "caov3", "parent")
names(exp.names) <- 1:8

get.pop.name <- function(bc) {
  exp.names[as.numeric(strsplit(bc, "-")[[1]][2])]
}

md.df <- data.frame(
  "traj.time" = traj$time,
  "experiment" = sapply(r14@cell.names, get.pop.name)
)

md.df <- md.df[order(md.df$traj),]
```

```{r}
caov3.cells <- md.df[md.df$experiment=="caov3",]
parent.cells <- md.df[md.df$experiment=="parent",]
s02.cells <- md.df[md.df$experiment=="s02",]
s03.cells <- md.df[md.df$experiment=="s03",]
step5.cells <- md.df[md.df$experiment=="r14_5u",]
step10.cells <- md.df[md.df$experiment=="r14_10u",]
step15E.cells <- md.df[md.df$experiment=="r14_15u_E",]
step15L.cells <- md.df[md.df$experiment=="r14_15u_L",]
```

```{r}
caov3.traj <- traj$time[rownames(caov3.cells)]
parent.traj <- traj$time[rownames(parent.cells)]
s02.traj <- traj$time[rownames(s02.cells)]
s03.traj <- traj$time[rownames(s03.cells)]
step5.traj <- traj$time[rownames(step5.cells)]
step10.traj <- traj$time[rownames(step10.cells)]
step15E.traj <- traj$time[rownames(step15E.cells)]
step15L.traj <- traj$time[rownames(step15L.cells)]
```

```{r}
r14.pops <- list(
  "caov3"=caov3.traj,
  "parent"=parent.traj,
  "s02"=s02.traj,
  "s03"=s03.traj,
  "step5"=step5.traj,
  "step10"=step10.traj,
  "step15E"=step15E.traj,
  "step15L"=step15L.traj
)
```

### Run per-cell enrichment

```{r}
enrichment <- chromasc::run_enrichment(r14, hallmarks, null_params)
```

### Visualize hallmark enrichments

```{r}
chromasc::show_populations_enrichment(enrichment, r14.pops, names(hallmarks))
```



